---
title: "HW5"
author: "Conor Dickson"
date: "2/12/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---


## Assign to the variable n_dims a single random integer between 3 and 10.
```{r}
n_dims <- sample(3:10, 1)
print(n_dims)
```
## Create a vector of consecutive integers from 1 to n_dims2.
```{r}
vector_1 <- c(1:(n_dims)^2)
print(vector_1)
```
## Use the sample function to randomly reshuffle these values.
```{r}
vector_2 <- sample(vector_1)
print(vector_2)
```
## create a square matrix with these elements.
```{r}
matrix_1 <- matrix(vector_2,nrow=n_dims,ncol=n_dims)

```
print out the matrix.
```{r}
print(matrix_1)
```

## find a function in r to transpose the matrix.
```{r}
trans.matrix1 <- t(matrix_1)

```

## print it out again and note how it has changed.
```{r}
print(trans.matrix1)
## rows have swapped with columns
```

## calculate the sum and the mean of the elements in the first row and the last row.
```{r}
sum <- c(mean(trans.matrix1[1,])+mean(trans.matrix1[4,]))
print(sum)
```

## read about the eigen() function and use it on your matrix
```{r}
eigen.stuff <- eigen(matrix_1)
print(eigen.stuff)
```

## look carefully at the elements of $values and $vectors. What kind of numbers are these?
```{r}
typeof(eigen.stuff)

```

dig in with the typeof() function to figure out their type.

## if have set your code up properly, you should be able to re-run it and create a matrix of different size because n_dims will change.
```{r}
print(matrix_1)
```

### Create a list with the following named elements:
## my_matrix, which is a 4 x 4 matrix filled with random uniform values
## my_logical which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.
##  my_letters, which is a 26-element vector of all the lower-case letters in random order.
```{r}
list <- list(my_matrix=matrix(runif(1:16),nrow=4,ncol=4),
             my_logical=c(runif(100)>0.5),
             my_letters=letters[sample(1:26)])

print(list)


```



## create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.
```{r}
list2 <- list(list[[1]][2,2],
              list[[2]][2],
              list[[3]][2])
print(list2)
```

## use the typeof() function to confirm the underlying data types of each component in this list
```{r}
typeof(list2[[1]])
typeof(list2[[2]])
typeof(list2[[3]])
```

## combine the underlying elements from the new list into a single atomic vector with the c() function.
```{r}
vector_3 <- c(list[[1]],list[[2]],list[[3]])
print(vector_3)
```

## what is the data type of this vector?
```{r}
typeof(vector_3)
```

### Create a data frame with two variables (= columns) and 26 cases (= rows).
## call the first variable my_unis and fill it with a random uniform value from 0 to 10
## call the second variable my_letters and fill it with 26 capital letters in random order.
```{r}
library(ggplot2)
data_frame1 <- data.frame(my_units=runif(n=26,min=0,max=10),
                          my_letters=sample(LETTERS))
print(data_frame1)
```


for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.
```{r}
library(base)
library(naniar)
print(data_frame1[,1])
data_frame1[,1] <- replace_with_na_all(replace=list(sample.int(x=data_frame1[,1],size=4)))

```

for the first variable, write a single line of R code to identify which rows have the missing values.

for the second variable, sort it in alphabetical order

calculate the column mean for the first variable.
