---
title: "HW8"
author: "Conor Dickson"
date: "3/5/2020"
output: html_document
---
```{r}
library(dplyr)
```

Examine the structure of the iris dataset. How many observations and variables are in the dataset?
```{r}
data(iris)
class(iris)
head(iris)
glimpse(iris)
```

Create a new data frame iris1 that contains only the species virginica and versicolor with sepal lengths longer than 6 cm and sepal widths longer than 2.5 cm. How many observations and variables are in the dataset?
```{r}
iris1 <- filter(iris, Species== c("virginica","versicolor") & Sepal.Length > 6.0 & Sepal.Width > 2.5)
glimpse(iris1)
```

Now, create a iris2 data frame from iris1 that contains only the columns for Species, Sepal.Length, and Sepal.Width. How many observations and variables are in the dataset?
```{r}
iris2 <- select(iris1,c(Species,Sepal.Length,Sepal.Width))
glimpse(iris2)
```

Create an iris3 data frame from iris2 that orders the observations from largest to smallest sepal length. Show the first 6 rows of this dataset.
```{r}
iris3 <- arrange(iris2, by= desc(Sepal.Length))
head(iris3)
```

Create an iris4 data frame from iris3 that creates a column with a sepal area (length * width) value for each observation. How many observations and variables are in the dataset?
```{r}
iris4 <- mutate(iris3,Sepal.Area=Sepal.Length*Sepal.Width)
glimpse(iris4)
```

Create iris5 that calculates the average sepal length, the average sepal width, and the sample size of the entire iris4 data frame and print iris5.
```{r}
iris5 <- summarize(iris4,avg.sepal.length=mean(Sepal.Length),avg.sepal.width=mean(Sepal.Width), sample.size=n())
print(iris5)
```

Finally, create iris6 that calculates the average sepal length, the average sepal width, and the sample size for each species of in the iris4 data frame and print iris6.
```{r}
iris4species <- group_by(iris4, Species)
iris6 <- summarize(iris4species,avg.sepal.length=mean(Sepal.Length),avg.sepal.width=mean(Sepal.Width), sample.size=n())
print(iris6)
```



